---
title: "Chapter 5: Problem 8"
output: html_notebook
---

Clear workspace, plots & console window
```{r}
# Clear workspace
rm(list = ls())
# Clear all plots
graphics.off()
# Clear Console
cat("\f")
```

Make sure that you have the directory set to the correct location for the data.
```{r}
# run this if it's your first time using it to install
# install.packages("rstudioapi")

# Make sure that you have the directory set to the correct location for the data.
library(rstudioapi) # load it
# The following line is for getting the path of your current open file
current_path <- getActiveDocumentContext()$path 
# The next line set the working directory to the relevant one:
setwd(dirname(current_path ))
# You can make sure you are in the right directory
print( getwd() )
```
We will now perform cross-validation on a simulated data set.
a) Generate a simulated data set as follows:
```{r}
set.seed(1)
x = rnorm(100)
y = x - 2*x^2 + rnorm(100)
```

In this data set, what is n and what is p? Write out the model used to generate the data in equation form.

n = 100, p = 2

b) Create a scatterplot of X against Y . Comment on what you find.
```{r}
plot(x,y)
```

c) Set a random seed, and then compute the LOOCV errors that result from fitting the following four models using least squares:

```{r}
df = data.frame(y,x)
error = matrix(data = 0, nrow = 100, ncol = 4)
for (i in 1:nrow(df))
{
  train = (1:nrow(df))[-i]
  lm.fit = list()#vector(mode = "list", length = 4)
  for (j in 1:4)
  {
    lm.fit[[j]] = lm(y~poly(x,j), data = df, subset = train)
    lm.pred = predict(lm.fit[[j]], df[-train,])
    error[i,j] = (df$y[-train] - lm.pred)
  }
}
```

