---
title: "Chapter 5: Problem 6"
output: html_notebook
---

Clear workspace, plots & console window
```{r}
# Clear workspace
rm(list = ls())
# Clear all plots
graphics.off()
# Clear Console
cat("\f")
```

Make sure that you have the directory set to the correct location for the data.
```{r}
# run this if it's your first time using it to install
# install.packages("rstudioapi")

# Make sure that you have the directory set to the correct location for the data.
library(rstudioapi) # load it
# The following line is for getting the path of your current open file
current_path <- getActiveDocumentContext()$path 
# The next line set the working directory to the relevant one:
setwd(dirname(current_path ))
# You can make sure you are in the right directory
print( getwd() )
```

We continue to consider the use of a logistic regression model to predict the probability of default using income and balance on the Default data set. In particular, we will now compute estimates for the standard errors of the income and balance logistic regression coefficients in two different ways: (1) using the bootstrap, and (2) using the standard formula for computing the standard errors in the glm() function. Do not forget to set a random seed before beginning your analysis.

a) Using the summary() and glm() functions, determine the estimated standard errors for the coefficients associated with income and balance in a multiple logistic regression model that uses both predictors.

```{r}
library(ISLR)
set.seed(1)
glm.fit=glm(default~income + balance, data = Default, family = binomial())
summary(glm.fit)
```
b) Write a function, boot.fn(), that takes as input the Default data set as well as an index of the observations, and that outputs the coefficient estimates for income and balance in the multiple logistic regression model.
```{r}
boot.fn=function (data ,index )
  + return (coef(glm(default~income + balance, data = data, subset =index, family = binomial)))
```
c) Use the boot() function together with your boot.fn() function to estimate the standard errors of the logistic regression coefficients for income and balance.
```{r}
boot(Default, boot.fn, 1000)
```
d) Comment on the estimated standard errors obtained using the glm() function and using your bootstrap function.

Standard errors are slightly reduced