---
title: "Chapter 4: Problem 13"
output: html_notebook
---

Clear workspace, plots & console window
```{r}
# Clear workspace
rm(list = ls())
# Clear all plots
graphics.off()
# Clear Console
cat("\f")
```

Make sure that you have the directory set to the correct location for the data.
```{r}
# run this if it's your first time using it to install
# install.packages("rstudioapi")

# Make sure that you have the directory set to the correct location for the data.
library(rstudioapi) # load it
# The following line is for getting the path of your current open file
current_path <- getActiveDocumentContext()$path 
# The next line set the working directory to the relevant one:
setwd(dirname(current_path ))
# You can make sure you are in the right directory
print( getwd() )
```
Chapter 4: Problem 13
IUsing the Boston data set, fit classification models in order to predict whether a given suburb has a crime rate above or below the median. Explore logistic regression, LDA, and KNN models using various subsets of the predictors. Describe your findings.
```{r}
library(MASS)
dim(Boston)
names(Boston)
summary(Boston)
```
Computing and adding median to dataframe
```{r}
crim.median = median(Boston$crim)
Boston$crim01 = ifelse(Boston$crim > crim.median, 1, 0)
Boston$crim01 = as.factor(Boston$crim01)
dim(Boston)
summary(Boston)
```
Computing correlation
```{r}
pairs(Boston, col = Boston$crim01)
dat2 <- data.frame(lapply(Boston, function(x) as.numeric(as.character(x))))
corr.d <- cor( dat2)
corr.d 
corr.d[ lower.tri( corr.d, diag = TRUE ) ] <- NA
library(corrplot)
corrplot( corr.d, type = "upper", diag = FALSE )
```

