---
title: "Chapter 4: Problem 12"
output: html_notebook
---

Clear workspace, plots & console window
```{r}
# Clear workspace
rm(list = ls())
# Clear all plots
graphics.off()
# Clear Console
cat("\f")
```
Make sure that you have the directory set to the correct location for the data.
```{r}
# run this if it's your first time using it to install
# install.packages("rstudioapi")

# Make sure that you have the directory set to the correct location for the data.
library(rstudioapi) # load it
# The following line is for getting the path of your current open file
current_path <- getActiveDocumentContext()$path 
# The next line set the working directory to the relevant one:
setwd(dirname(current_path ))
# You can make sure you are in the right directory
print( getwd() )
```
**Make sure that you have the directory set to the correct location for the data.
```{r}
# run this if it's your first time using it to install
# install.packages("rstudioapi")

# Make sure that you have the directory set to the correct location for the data.
library(rstudioapi) # load it
# The following line is for getting the path of your current open file
current_path <- getActiveDocumentContext()$path 
# The next line set the working directory to the relevant one:
setwd(dirname(current_path ))
# You can make sure you are in the right directory
print( getwd() )
```
**Make sure that you have the directory set to the correct location for the data.
```{r}
# run this if it's your first time using it to install
# install.packages("rstudioapi")

# Make sure that you have the directory set to the correct location for the data.
library(rstudioapi) # load it
# The following line is for getting the path of your current open file
current_path <- getActiveDocumentContext()$path 
# The next line set the working directory to the relevant one:
setwd(dirname(current_path ))
# You can make sure you are in the right directory
print( getwd() )
```
**This problem involves writing functions.**
a) Write a function, Power() , that prints out the result of raising 2 to the 3rd power. In other words, your function should compute $2^3$ and print out the results.
*Hint: Recall that x^a raises x to the power a . Use the print() function to output the result.*

```{r}
Power = function()
{
  cat(2^3)
}

Power()
```
b) Create a new function, Power2() , that allows you to pass any two numbers, x and a, and prints out the value of $x^a$ .
```{r}
Power2 = function(x,a)
{
  cat(x^a)
  cat('\n')
}

Power2(3,8)
```
c) Using the Power2() function that you just wrote, compute $10^3$, $8^17$, and $131^3$ .
```{r}
Power2(10,3)
Power2(8,17)
Power2(131,3)
```
d) Now create a new function, Power3() , that actually returns the result $x^a$ as an R object, rather than simply printing it to the screen. That is, if you store the value $x^a$ in an object called result within your function, then you can simply return() this result, using the following line:
```{r}
Power3 = function(x,a=2)
{
  return(x^a)
}
```

e)Now using the Power3() function, create a plot of $f(x) = x^2$. The x-axis should display a range of integers from 1 to 10, and the y-axis should display $x^2$. Label the axes appropriately, and use an appropriate title for the figure. Consider displaying either the x-axis, the y-axis, or both on the log-scale. You can do this by using log=‘‘x’’ , log=‘‘y’’ , or log=‘‘xy’’ as arguments to the plot() function.
```{r}
par(mfrow = c(2, 2))
for(ll in c("", "x", "y", "xy"))
   curve(Power3, 1, 10, log = ll, sub = paste0("log = '", ll, "'"))

```
f) Create a function, PlotPower() , that allows you to create a plot of x against x^a for a fixed a and for a range of values of x.
```{r}
PlotPower = function(x, a=3)
{
  y = Power3(x,a)
  plot(x,y)
}
PlotPower(1:10,3)
```

