---
title: "Chapter 4: Problem 11"
output: html_notebook
---

Clear workspace, plots & console window
```{r}
# Clear workspace
rm(list = ls())
# Clear all plots
graphics.off()
# Clear Console
cat("\f")
```

Make sure that you have the directory set to the correct location for the data.
```{r}
# run this if it's your first time using it to install
# install.packages("rstudioapi")

# Make sure that you have the directory set to the correct location for the data.
library(rstudioapi) # load it
# The following line is for getting the path of your current open file
current_path <- getActiveDocumentContext()$path 
# The next line set the working directory to the relevant one:
setwd(dirname(current_path ))
# You can make sure you are in the right directory
print( getwd() )
```
Chapter 4: Problem 11
In this problem, you will develop a model to predict whether a given car gets high or low gas mileage based on the Auto data set.
```{r}
library(ISLR)
#attach(Weekly)
dim(Auto)
names(Auto)
summary(Auto)
```
a) Create a binary variable, mpg01 , that contains a 1 if mpg contains a value above its median, and a 0 if mpg contains a value below its median. You can compute the median using the median() function. Note you may find it helpful to use the data.frame() function to create a single data set containing both mpg01 and the other Auto variables.
```{r}
mpg.median = median(Auto$mpg)
Auto$mpg01 = ifelse(Auto$mpg > mpg.median, 1, 0)
Auto$mpg01 = as.factor(Auto$mpg01)
dim(Auto)
summary(Auto)
```
b) Explore the data graphically in order to investigate the association between mpg01 and the other features. Which of the other features seem most likely to be useful in predicting mpg01 ? Scatterplots and boxplots may be useful tools to answer this question. Describe your findings.
```{r}
pairs(Auto)
```

```{r}
boxplot(Auto)
```

