---
title: "Chapter 4: Problem 11"
output: html_notebook
---

Clear workspace, plots & console window
```{r}
# Clear workspace
rm(list = ls())
# Clear all plots
graphics.off()
# Clear Console
cat("\f")
```

Make sure that you have the directory set to the correct location for the data.
```{r}
# run this if it's your first time using it to install
# install.packages("rstudioapi")

# Make sure that you have the directory set to the correct location for the data.
library(rstudioapi) # load it
# The following line is for getting the path of your current open file
current_path <- getActiveDocumentContext()$path 
# The next line set the working directory to the relevant one:
setwd(dirname(current_path ))
# You can make sure you are in the right directory
print( getwd() )
```
Chapter 4: Problem 11
In this problem, you will develop a model to predict whether a given car gets high or low gas mileage based on the Auto data set.
```{r}
library(ISLR)
#attach(Weekly)
dim(Auto)
names(Auto)
summary(Auto)
```
a) Create a binary variable, mpg01 , that contains a 1 if mpg contains a value above its median, and a 0 if mpg contains a value below its median. You can compute the median using the median() function. Note you may find it helpful to use the data.frame() function to create a single data set containing both mpg01 and the other Auto variables.
```{r}
mpg.median = median(Auto$mpg)
Auto$mpg01 = ifelse(Auto$mpg > mpg.median, 1, 0)
Auto$mpg01 = as.factor(Auto$mpg01)
dim(Auto)
summary(Auto)
```
b) Explore the data graphically in order to investigate the association between mpg01 and the other features. Which of the other features seem most likely to be useful in predicting mpg01 ? Scatterplots and boxplots may be useful tools to answer this question. Describe your findings.
```{r}
pairs(Auto, col = Auto$mpg01)
dat2 <- data.frame(lapply(Auto[which(!names(Auto)=="name")], function(x) as.numeric(as.character(x))))
corr.d <- cor( dat2)
corr.d 
corr.d[ lower.tri( corr.d, diag = TRUE ) ] <- NA
library(corrplot)
corrplot( corr.d, type = "upper", diag = FALSE )
```
c) Split the data into a training set and a test set.
```{r}
# Spliting data with mpg > median
Auto.mpg1 = Auto[which(Auto$mpg01 == 1),]
# Spliting data with mpg < median
Auto.mpg0 = Auto[which(Auto$mpg01 == 0),]
# Permute data for mpg > median
x = sample(1:nrow(Auto.mpg1))
Auto.mpg1 = Auto.mpg1[x,]
# Permute data for mpg < median
x = sample(1:nrow(Auto.mpg0))
Auto.mpg0 = Auto.mpg0[x,]
# training data for mpg > median
Auto.mpg1.train = Auto.mpg1[1:floor(nrow(Auto.mpg1)*2/3),]
# test data for mpg > median
Auto.mpg1.test = Auto.mpg1[ceiling(nrow(Auto.mpg1)*2/3):nrow(Auto.mpg1),]
# training data for mpg < median
Auto.mpg0.train = Auto.mpg0[1:floor(nrow(Auto.mpg0)*2/3),]
# test data for mpg < median
Auto.mpg0.test = Auto.mpg0[ceiling(nrow(Auto.mpg0)*2/3):nrow(Auto.mpg0),]
# Overall training data
Auto.train = rbind(Auto.mpg0.train, Auto.mpg1.train)
# Overall test data
Auto.test = rbind(Auto.mpg0.test, Auto.mpg1.test)
# Delete temperary variables from workspace
rm(Auto.mpg1, Auto.mpg0, Auto.mpg1.train, Auto.mpg1.test, Auto.mpg0.train, Auto.mpg0.test, x)
```
d) Perform LDA on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained?
```{r}

library(MASS)
lda.fit = lda(mpg01~.-name, data=Auto.train)
lda.fit
plot(lda.fit)
lda.pred = predict(lda.fit, Auto.test)
names(lda.pred)
lda.class=lda.pred$class
table(lda.class, Auto.test$mpg01)
mean(lda.class!=Auto.test$mpg01)
```
e) Perform QDA on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained?
```{r}
qda.fit = qda(mpg01~.-name, data=Auto.train)
qda.fit
#plot(qda.fit)
qda.pred = predict(qda.fit, Auto.test)
names(qda.pred)
qda.class=qda.pred$class
table(qda.class, Auto.test$mpg01)
mean(qda.class!=Auto.test$mpg01)
```
f) Perform logistic regression on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained?
```{r}
glm.fits=glm(mpg01~weight, data = Auto.train, family = binomial)
summary(glm.fits)
glm.probs=predict(glm.fits, Auto.test, type="response")
glm.pred = rep(0, nrow(Auto.test))
glm.pred[glm.probs>0.5] = 1
table(glm.pred,  Auto.test$mpg01)
mean(glm.pred!= Auto.test$mpg01)
```
g) Perform KNN on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained? Which value of K seems to perform the best on this data set?
```{r}
train.X = scale(Auto.train$acceleration)
test.X = scale(Auto.test$acceleration)
train.Y = Auto.train$mpg01
test.Y = Auto.test$mpg01
set.seed(1)
knn.pred = knn(train.X, test.X, train.Y, k=3)
mean(test.Y!=knn.pred)
mean(test.Y!="No")
```

